---
title: "group4"
author: "Siqi Wang, Yubin Lyu, Liting Wang, Han Xu, JIahao Mao"
output: pdf_document
---

```{r setup, include=FALSE}
options(scipen = 1, digits = 2) #set to two decimal
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA)
```


```{r libraries}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(moderndive)
library(gapminder)
library(skimr)
library(corrplot)
library(ggcorrplot)
library(GGally)
library(janitor)
library(gridExtra)
library(MASS)
library(BMA)
library(moments)
library(kableExtra)
```

# Introduction

The Family Income and Expenditure Survey (FIES) is a significant source of data for understanding the wellbeing of households in Philippines. It provides valuable information on family income and expenditure, which can be used to investigate various research questions related to household characteristics.

In this analysis, we are interested in identifying which household-related factors influence the size of a household. Using Generalized Linear Model (GLM), we will explore the datasets obtained from the FIES survey for XII - SOCCSKSARGEN region in Philippines. The results of our analysis could help the government to make informed decisions related to household policies and other related matters.


## Data Processing
```{r data, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
household <- read.csv("dataset4.csv") 
household$Electricity <- as.factor(household$Electricity)
household$Household.Head.Sex <- as.factor(household$Household.Head.Sex)
household$Type.of.Household <- as.factor(household$Type.of.Household)

```

```{r rename, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}

colnames(household)[1]<-"Income"             # original name "Total.Household.Income" 
colnames(household)[3]<-"FoodExp"            # original name "Total.Food.Expenditure"   
colnames(household)[4]<-"Householder_Sex"    # original name "Household.Head.Sex"
colnames(household)[5]<-"Householder_Age"    # original name "Household.Head.Age"
colnames(household)[6]<-"Household_Type"     # original name "Type.of.Household"
colnames(household)[7]<-"Number_Members"     # original name "Total.Number.of.Family.members"
colnames(household)[8]<-"Floorarea"          # original name "House.Floor.Area" 
colnames(household)[10]<-"Number_bedrooms"   # original name "Number.of.bedrooms"
```

## Data Summary
```{r skim, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
household_cat <- household %>%
   dplyr::select("Electricity","Householder_Sex","Household_Type")
summary(household_cat)

household_num <- household[, sapply(household, is.numeric)]
my_skim <- skim_with(base = sfl(n = length))
household_num %>%
  my_skim() %>%
  transmute(Variable=skim_variable, n=n, 
            Mean = format(signif(numeric.mean, 3), scientific = TRUE, digits = 2), 
            SD = format(signif(numeric.sd, 3), scientific = TRUE, digits = 2),
            Min= format(signif(numeric.p0, 3), scientific = TRUE, digits = 2), 
            Median=format(signif(numeric.p50, 3), scientific = TRUE, digits = 2),  
            Max=format(signif(numeric.p100, 3), scientific = TRUE, digits = 2), 
            IQR = format(signif(numeric.p75-numeric.p50, 3), scientific = TRUE, digits = 2) ) %>%
   kable(caption = '\\label{tab:summarybyskim} Summary statistics of variables',
         booktabs = TRUE, linesep = "", digits = 2) %>%
   kable_styling(font_size = 10, latex_options = "hold_position")

```



# Distribution Check

test if the distribution of y is poisson dist

```{r pcheck, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}

hist(household$Number_Members)

```
 
check the skewness and kurtosis results

```{r nbcheck, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}

skewness(household$Number_Members)
kurtosis(household$Number_Members)

```
Based on the skewness and kurtosis results, we can determine that the distribution of "y" does not conform to the assumption of a strict Poisson distribution. 
Specifically, skewness values greater than 1 indicate that the data distribution is right-skewed, and kurtosis values greater than 3 indicate that the data distribution is sharper than the Poisson distribution.
In such cases, a Negative Binomial Distribution (NBD) regression model may be considered, as it can be fitted when a Poisson regression model is not up to the task. 


# Correlation Matrix and GGpairs

```{r correlation, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}

cor_matrix <- cor(household_num) %>% #create the correlation matrix of variables
  round(2)
cor_matrix

corrplot <- corrplot(cor_matrix, method = "color", addCoef.col = "gray",type = "upper",)
ggpairs(household_num)

```
# Boxplots of Variables

```{r boxplot, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}

ggplot(data = household, mapping = aes(x = Householder_Sex, y = Number_Members)) +
  geom_boxplot(aes(fill =Householder_Sex ) )+
  labs(x = "Householder_Sex", y = "Number_Members",
       title = "gender")  + 
  scale_x_discrete(labels = c("female","male"))

ggplot(data = household, mapping = aes(x = Household_Type, y = Number_Members)) +
  geom_boxplot(aes(fill = Household_Type))+
  labs(x = "Household_Type", y = "Number_Members",title = "type") +
  scale_x_discrete(labels = c("Extended", "Single", "Two or More"))+
  theme(legend.position = "bottom")

ggplot(data = household, mapping = aes(x = Electricity, y = Number_Members)) +
  geom_boxplot(aes(fill =Electricity ) )+
  labs(x = "Electricity", y = "Number_Members",
       title = "Electricity")  + 
  scale_x_discrete(labels = c("no","yes"))

```
# Histogram of Variables

```{r histogram, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}

household %>% 
  tabyl(Number_Members, Householder_Sex) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() # To show original counts

# barplot of Number_Members and Householder_Sex
ggplot(household, aes(x= Number_Members,  y = ..prop.., group=Householder_Sex, fill=Householder_Sex)) + 
  geom_bar(position="dodge", stat="count") +
  labs(y = "Proportion")

# barplot of Number_Members and Electricity 
ggplot(household, aes(x= Number_Members,  y = ..prop.., group=Electricity, fill=Electricity)) + 
  geom_bar(position="dodge", stat="count") +
  labs(y = "Proportion")

# barplot of Number_Members and Household_Type
ggplot(household, aes(x= Number_Members,  y = ..prop.., group=Household_Type, fill=Household_Type)) + 
  geom_bar(position="dodge", stat="count") +
  labs(y = "Proportion") +
  theme(legend.position = "bottom")

```

# Model Fitting

```{r poissson, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}

#model with all variables
m1 <- glm(formula = Number_Members ~ Income + FoodExp + Householder_Sex + 
            Householder_Age + Household_Type + Floorarea +
            House.Age + Number_bedrooms + Electricity, 
          family = poisson(link = "log"), data = household)

#model with log transformation of Income and FoodExp
m2 <- glm(formula = Number_Members ~ log(Income) + log(FoodExp) + 
            Householder_Age + Floorarea +
            House.Age + Number_bedrooms, 
          family = poisson(link = "log"), data = household)

summary(m1)
summary(m2)

```

## Use BIC to do variable selection

```{r BIC, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}

output <- bic.glm(Number_Members ~ Income + FoodExp + Householder_Sex + 
                    Householder_Age + Household_Type + Floorarea +
                    House.Age + Number_bedrooms + Electricity, 
                   glm.family = "poisson" , data = household)
summary(output)

m3 <- glm(Number_Members ~ Income + FoodExp + Householder_Sex + 
                    Householder_Age + Household_Type +
                    House.Age + Number_bedrooms + Electricity, 
                   family = "poisson" , data = household)
summary(m3)
    
```

## Negative Binomial Distribution
```{r NB, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}

m4 <- glm.nb(formula = Number_Members ~ Income + FoodExp + Householder_Sex + 
            Householder_Age + Household_Type + Floorarea +
            House.Age + Number_bedrooms + Electricity, data = household)
summary(m4)


```

# Deviance plots

```{r deviance, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
resp <- resid(m1, type = "pearson") 
resd <- resid(m1, type = "deviance")
r1<- ggplot(m1, aes(sample = resp)) + geom_point(stat = "qq", color = "#7fc97f") + ylab("Pearson residuals")
r2<- ggplot(m1, aes(sample = resd)) + geom_point(stat = "qq", color = "#7fc97f") + ylab("Deviance residuals")
r3<- ggplot(m1, aes(x = predict(m1, type="link"), y =resd))+ geom_point(col = "#7fc97f") +
  ylab("Deviance residuals") + xlab("Linear predictor") 
grid.arrange(r1, r2, r3, nrow = 1)

resp2 <- resid(m2, type = "pearson") 
resd2 <- resid(m2, type = "deviance")
r4<- ggplot(m2, aes(sample = resp2)) + geom_point(stat = "qq", color = "#7fc97f") + ylab("Pearson residuals")
r5<- ggplot(m2, aes(sample = resd2)) + geom_point(stat = "qq", color = "#7fc97f") + ylab("Deviance residuals")
r6<- ggplot(m2, aes(x = predict(m2, type="link"), y =resd2))+ geom_point(col = "#7fc97f") +
  ylab("Deviance residuals") + xlab("Linear predictor") 
grid.arrange(r4, r5, r6, nrow = 1)
```

# Model Evaluation

```{r model_evaluation, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
# Poisson model
c(m1$deviance, m1$aic)
# poission model with log transformation
c(m2$deviance, m2$aic)
# BIC model
c(m3$deviance, m3$aic)
# Negative binomial model
c(m4$deviance, m4$aic)
```

# Goodness-of-fit test

```{r GOF, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}

chisq <- with(m1, sum((household$Number_Members- fitted.values)^2/fitted.values))
df <- with(m1, df.residual)
p <- with(m1, pchisq(chisq, df, lower.tail = FALSE))
cat("Chi-square test statistic = ", chisq, "\n")
cat("df = ", df, "\n")
cat("p-value = ", p, "\n")


# The coef() function obtains the coefficients of the model.
# The confint() function obtains the confidence interval of the model coefficients.

exp(cbind(OR = coef(m1), confint(m1)))

#The OR value is equal to the regression coefficient of the indexed variable. 
#Specifically, if the coefficient of an explanatory variable in the regression model is b, its corresponding OR value is exp(b)


```